---
title: "MLE_Normal_distribution"
author: "Fallaw Sowell"
date: "August 29, 2018"
output:
  pdf_document: null
  number_sections: yes
  html_document:
    df_print: paged
  toc: yes
  toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Program Description

This program perform maximum likelihood estimation
for a normal likelihood function.  The log-likelihood
function is maximized numerically.
Analytical and numerical second derivatives are compared.

```{r cleanup}
rm(list = ls())
```

## Load Packages

Load Packages and Set Options

```{r packages}
if (!require('openxlsx'))
  install.packages('openxlsx')
library('openxlsx')

if (!require('ggplot2'))
  install.packages('ggplot2')
library('ggplot2')
```

## User Defined Functions


This is the code that evaluates the normal log-likelihood function

```{r Normal_log_likelihood}
Normal_log_like <- function(the_parameters, the_data) {
  # Computes the log-likelihood function for a normal distribution
  #
  # Args:
  #   the_parameters: First the mean and then the standard deviation.
  #   the_data: A column matrix of the data.
  #
  # Returns:
  #   The log-likelihood function evaluated at the parameters.
  mu <- the_parameters[1]
  sigma  <- the_parameters[2]
  x <- the_data
  N <- length(x)
  
  log_like <- (-(N / 2) * log(2 * pi)
               - N * log(sigma)
               - (1 / (2 * sigma * sigma)) * (sum(x ^ 2) - 2 * mu * sum(x) + N * mu * mu))
  return(-log_like)
}
```

Next is the user defined function that calculates the second derivative of the 
normal log-likelihood function.

```{r Hessian_function}
Hessian_Normal_log_like <- function(the_parameters, the_data) {
  # Computes the second derivative of a normal log-likelihood function
  #
  # Args:
  #   the_parameters: First the mean and then the standard deviation.
  #   the_data: A column matrix of the data.
  #
  # Returns:
  #   The 2x2 matrix of second derivatives evaluated at the given parameters.
  mu <- the_parameters[1]
  sigma  <- the_parameters[2]
  x <- the_data
  N <- length(x)
  D_mu2 <- -N / (sigma ^ 2)
  D_sigma2 <- (N / (sigma ^ 2)
               - (3 / (sigma ^ 4)) * (sum(x ^ 2) - 2 * mu * sum(x) + N * mu * mu))
  D_mu_sigma <- -(2 / (sigma ^ 3)) * (sum(x) - N * mu)
  H <- matrix(c(D_mu2, D_mu_sigma, D_mu_sigma, D_sigma2), ncol = 2)
  return(H)
}
```


## Data Summary
```{r read_data}
Normal_df <-
  read.xlsx(xlsxFile = "MLE_data_set_normal.xlsx", sheet = "Normal")
X <- as.matrix(Normal_df[1])
N <- length(X)
```


It is nice to learn about your data with a plot and summary statistics

```{r summary_and_plot, message=FALSE, warning=FALSE}
summary(Normal_df)

g <- ggplot(Normal_df, aes(x = X)) + geom_histogram()
g
```

## Estimation and Confidence Intervals

Use the R function optim() to minimize the log-likelihood function.  We must set the start values for the numerical hill-climber.  

```{r Maximize_the_log_like}
start_values = c(1.2, 1.8)

optout_normal <-
  optim(
    start_values,
    Normal_log_like,
    control = list(
      maxit = 500,
      abstol = .00000001,
      reltol = .00000001
    ),
    the_data = X
  )
optout_normal
```


```{r Calculate_Covarinace}
the_covariance <-
  solve(-Hessian_Normal_log_like(optout_normal$par , X) / N) / N
the_covariance
```

Now calculate and print the 95% confidence intervals for the parameters.  

```{r confidence_intervals}
print(c(
  optout_normal$par[1] - 2 * sqrt(the_covariance[1, 1]),
  optout_normal$par[1] + 2 * sqrt(the_covariance[1, 1])
))

print(c(
  optout_normal$par[2] - 2 * sqrt(the_covariance[2, 2]),
  optout_normal$par[2] + 2 * sqrt(the_covariance[2, 2])
))
```

## Compare Analytical and Numerical Second Derivatives

First print the analytical second derivatives.  Then, using a delta of .001, numerically approximate the second derivatives.  The numerical values are then
displayed so they can be compared to the analytical second derivatives. 

```{r Numerical_Second_derivative}

Hessian_Normal_log_like(optout_normal$par , X)

delta <- .001
Num_H11 <- -(
  Normal_log_like(optout_normal$par - c(delta, 0), X)
  - 2 * Normal_log_like(optout_normal$par, X)
  + Normal_log_like(optout_normal$par + c(delta, 0), X)
) / (delta * delta)

Num_H22 = -(
  Normal_log_like(optout_normal$par - c(0, delta), X)
  - 2 * Normal_log_like(optout_normal$par, X)
  + Normal_log_like(optout_normal$par + c(0, delta), X)
) / (delta * delta)

Num_H11
Num_H22
```


