---
title: "Multiclass Logistic Regression Insurance"
author: "Fallaw Sowell"
date: "Sept 2, 2018"
output:
  pdf_document: null
  number_sections: yes
  html_document:
    df_print: paged
  toc: yes
  toc_depth: 4
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\tableofcontents 
\newpage

## Multiclass Logistic Regression Example 

This file was created with R markdown and Knit.  This demonstrates how to read in a data file, obtain summary statistics, run a multiclass logistic regression model and finally report confusion matrix.    

### Load the needed libraries

```{r load libraries, message=FALSE, warning=FALSE}

rm(list=ls())

if (!require('openxlsx')) install.packages('openxlsx')
library('openxlsx')

if (!require('ggplot2')) install.packages('ggplot2')
library('ggplot2')

if (!require('dplyr')) install.packages('dplyr')
library('dplyr')

if (!require('data.table')) install.packages('data.table')
library('data.table')

if (!require('plotmo')) install.packages('plotmo')
library('plotmo')

if (!require('nnet')) install.packages('nnet')
library('nnet')
```

### Load the data into R

First, read the data from the excel file and save as a data.frame.   This excel file needs to be in the R working directory.   Alternatively, you would need to include the path the excel file.  

This data is from an insurance company that is trying to understand why different charges occur for different customers.   We have the annual charges for the policy.  We have the sex, age, BMI, smoking status, and region of residence for the person who owns the policy.  We also have the number of children covered by the policy.   

```{r Read the data, echo=TRUE }
Insurance_df <- read.xlsx(xlsxFile = "Insurance.xlsx")
```

###  Get familiar with your data, variables, size, etc. 

It is good practice to review the data by looking at its structure.  

```{r Look at the structure of the data.frame}
str(Insurance_df)
```

###  Transform the data as needed

Note that the sex, smoker, and regions variables are listed as characters variables.  For us to use them in our analysis we need to convert them to factor variables. 

```{r converte character variables to factors}
Insurance_df$sex = as.factor(Insurance_df$sex)
Insurance_df$smoker = as.factor(Insurance_df$smoker)
Insurance_df$region = as.factor(Insurance_df$region)

#
#  split into three categories 
#

Insurance_df$category[Insurance_df$age > 2.5 + Insurance_df$charges /   300] <- 'low'
Insurance_df$category[
   Insurance_df$age < 2.5 + Insurance_df$charges / 300 &
   Insurance_df$age > -120 + Insurance_df$charges / 200] <- 'medium'
Insurance_df$category[Insurance_df$age < -120 + Insurance_df$charges / 200] <- 'high'

Insurance_df$category = as.factor(Insurance_df$category)

Insurance_df <-mutate(Insurance_df,charges = log(charges)) 
Insurance_df <-mutate(Insurance_df, age2 = age*age) 
```

### Learn about the data set with Summary statistics and plots

Get familiar with the data. Look at some summary statistics for the variables.

```{r summary statistics}
summary(Insurance_df)
```

We are interested in the relationships in the entire data set. A nice function that achieve this is "GGally::ggpairs" from the GGally and ggplot2 packages. 

```{r GGally_ggpairs, message=FALSE, warning=FALSE}
GGally::ggpairs(Insurance_df)
```

### Fit a multinomial Logistic Regression model to explain the class of charges

I will normalize with the `high` class.

```{r fit_multinomial_logit}

Insurance_df$category2 <- relevel(Insurance_df$category, ref = 'high')
#multi1 = multinom(category2 ~ ., data=Insurance_df)
multi1 = multinom(category2 ~ bmi+smoker, data=Insurance_df)

#
```

The output from the multinom function can be viewed with the summary command. 
```{r summarize multinom output}
summary(multi1)
```

## Diagnostic information is the confusion matrix

```{r summarize multinom fit}
predicted_scores <- predict (multi1, Insurance_df, "probs")
head(predicted_scores)

predicted_class <- predict (multi1, Insurance_df)
head(predicted_class)

table(predicted_class, Insurance_df$category)
```
