---
title: "Stats Prob set 2"
author: "Steven M. Barnard"
date: "September 23, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
## Problem Set 2 - Maximum Likelihood Estimation

```{r cleanup}
rm(list = ls())
# command to clean workspace out
```

## Load Packages

Load Packages and Set Options

```{r packages}
if (!require('openxlsx'))
  install.packages('openxlsx')
library('openxlsx')

if (!require('ggplot2'))
  install.packages('ggplot2')
library('ggplot2')
```

## Load Dataset

```{r load_data}
gamma_df <-
  read.xlsx(xlsxFile = "MLE_data_set_gamma.xlsx")
X <- as.matrix(gamma_df[1])
N <- length(X)
summary(gamma_df)
```

```{r summary_and_plot, message=FALSE, warning=FALSE}
g <- ggplot(gamma_df, aes(x = X)) + geom_histogram()
g
```
There appears to be some heteroskedasticity towards the mean with a large right sided tail. This is indicative of a gamma distribution.

## Analytical solutions
I have referenced the write-up provided by professor Sowell. My user defined functions should emulate the analytical solutions derived from the write-up.

## User Defined Functions

User defined functions to evaluate the gamma function.

```{r gamma_log_likelihood}
gamma_log_like <- function(the_parameters, the_data) {
  # Computes the log-likelihood function for a normal distribution
  #
  # Args:
  #   the_parameters: First the shape and then the scale.
  #   the_data: A column matrix of the data.
  #
  # Returns:
  #   The log-likelihood function evaluated at the parameters.
  alpha <- the_parameters[1] # 1st parameter - shape
  theta  <- the_parameters[2] # 2nd parameter - scale
  x <- the_data
  N <- length(x)
  
  log_like <- ((alpha-1) * sum(log(x)) - (1/theta) * sum(x) - alpha*N*log(theta)
               - N * lgamma(alpha))
  return(-log_like)
}
```

Next is the user defined function that calculates the second derivative of the 
gamma log-likelihood function.

The Trigamma function should return the 2x2 matrix; the covariance matrix for the two estimates.

```{r Trigamma}
Trigamma_log_like <- function(the_parameters, the_data) {
  # Computes the second derivative of a gamma log-likelihood function
  #
  # Args:
  #   the_parameters: First the shape and then the scale.
  #   the_data: A column matrix of the data.
  #
  # Returns:
  #   The 2x2 matrix of second derivatives evaluated at the given parameters.
  alpha <- the_parameters[1]
  theta  <- the_parameters[2]
  x <- the_data
  N <- length(x)
  d2_alpha <- -N * trigamma(alpha)
  d2_alpha_theta <- -N/theta
  d2_theta <- (-2/(theta * theta * theta)) * sum(x) + (alpha*N/(theta * theta)) 
  H <- matrix(c(d2_alpha,d2_alpha_theta ,d2_alpha_theta ,d2_theta ), ncol = 2)
  return(H)
}
```


## Estimation and Confidence Intervals

Use the R function optim() to minimize the log-likelihood function.  We must set the start values for the numerical hill-climber.  

```{r Maximize_the_log_like}
start_values = c(1.5, 1.8)

optout_normal <-
  optim(
    start_values,
    gamma_log_like,
    control = list(
      maxit = 500,
      abstol = .00000001,
      reltol = .00000001
    ),
    the_data = X
  )
optout_normal
```
The returned parameters are an alpha of ~1.87 and a theta of ~3.14. These our our maximum likelihood estimates.

The value of the log-likelihood function is ~668.81.


```{r Calculate_Covarinace}
the_covariance <-
  solve(-Trigamma_log_like(optout_normal$par , X) / N) / N
the_covariance
```
The estimate for the covariance of the shape and scale are returned as a 2x2 matrix.


Now we can calculate and print the 95% confidence intervals for the parameters.  

```{r confidence_interval for shape/alpha}
print(c(
  optout_normal$par[1] - 2 * sqrt(the_covariance[1, 1]),
  optout_normal$par[1] + 2 * sqrt(the_covariance[1, 1])
))
```
The 95% confidence interval for the alpha is ~[1.56, 2.18].
```{r confidence_interval for scale/theta}
print(c(
  optout_normal$par[2] - 2 * sqrt(the_covariance[2, 2]),
  optout_normal$par[2] + 2 * sqrt(the_covariance[2, 2])
))
```
The 95% confidence interval for the theta is ~[2.55, 3.74].

## Compare Analytical and Numerical Second Derivatives

I ran the Trigamma function then created a function to pass delta values to the numerical derivative example and stored them in a new df.

```{r 2nd deriv+Numerical_Second_derivative}

Trigamma_log_like(optout_normal$par , X)


Num_2nd <- function(delta_value) {
  # Computes the second derivative numerically with respect to alpha over the delta values.
  Num_T <- -(
  gamma_log_like(optout_normal$par - c(delta_value, 0), X)
  - 2 * gamma_log_like(optout_normal$par, X)
  + gamma_log_like(optout_normal$par + c(delta_value, 0), X)
) / (delta_value * delta_value)
  return(Num_T)
}

Yt <- c(Num_2nd(1.0e-1),Num_2nd(1.0e-2),Num_2nd(1.0e-3),Num_2nd(1.0e-4),Num_2nd(1.0e-5),Num_2nd(1.0e-6))

Xt<- c(1.0e-1,1.0e-2,1.0e-3,1.0e-4,1.0e-5,1.0e-6)

#Yt <- c(Num_2nd(1.0e-1),Num_2nd(1.0e-2),Num_2nd(1.0e-3),Num_2nd(1.0e-4),Num_2nd(1.0e-5),Num_2nd(1.0e-6),Num_2nd(1.0e-7),Num_2nd(1.0e-8),Num_2nd(1.0e-9),Num_2nd(1.0e-10),Num_2nd(1.0e-11),Num_2nd(1.0e-12))

#Xt<- c(1.0e-1,1.0e-2,1.0e-3,1.0e-4,1.0e-5,1.0e-6,1.0e-7,1.0e-8,1.0e-9,1.0e-10,1.0e-11,1.0e-12)

# I chose not to include all of the 12 entries so that the data could be better visualized in the plot, so that the 2nd derivative can be seen appraoching -175.50567 rather than having a visual that would have been less informative.

NumT_df <- data.frame(x=Xt,y=Yt)

NumT_df
```

```{r NumT_df plot}

ggplot(data = NumT_df) +
  geom_point(mapping = aes(log(x), y))
```

The graphical representation shows the max value approaching ~ -175.5057 before trailing off again at the peripheral x-values. This is nearly identical to our numerical estimate of the second derivative.