---
title: "Business Regression"
author: "Steven M. Barnard"
date: "September 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, include=FALSE}
if (!require('openxlsx')) install.packages('openxlsx')
library('openxlsx')
if (!require('ggplot2')) install.packages('ggplot2')
library('ggplot2')
if (!require('dplyr')) install.packages('dplyr')
library('dplyr')
if (!require('GGally')) install.packages('GGally')
library('GGally')
if (!require('plotmo')) install.packages('plotmo')
library('plotmo')
if (!require('data.table')) install.packages('data.table')
library('data.table')
if (!require('glmnet')) install.packages('glmnet')
library('glmnet')
```

Goal of this is to create a linear regression that helps determine what variables are meaningful for each product(watches) in total and in their respective markets.

## Game tips/info
Company is based in the US.

We produce 2 brands of watches.

We sell them in 6 countries: 
- US
- China
- UK
- Germany
- Japan
- Mexico

What we need to monitor:
- competition
- cust. pref.
- how we monitor our actions and our competitors

# Factories:

Each makes only 1 product (can be 1 or 2 but not both)
You can increase/decrease facotry size.
Both can be in the same country.
Coutry location affects quality.

You can produce and market 1 type of watch if you want. You dont need to produce and market 2 products.

# Product design:

Product 1: 
- more price sensitive
- low product loyalty
- will buy from competitors if you 'stock out'

Product 2:
- premium product
- faced with 'stock out' customers buy nothing and wait.

Product 2 and 1 are independent. (models need otbe run separately)

Customer preferences are independent per country. (separate model set up for each country)

# Market Design:
Customer preference generalizations will hold true.
Affluent countries have less price sensitive consumers.

Variables that affect decisions:
- price
- unit sales
- demand
- tariff rates
- transport rates/capacity
- green spend
- market spend (consualtants and actual marketing)
- currency exchange
- mfg cost

** early payment discount 0-2%

** Customer preferences are more important than factory location



# Loading the data
First the data base needs to be completed by gathering the necessary data from the current excel spreadsheet via the tabulate sheet and creating a new excel file to act as the database.

This database will have to be updated every week.

```{r}
#getwd() to check working directory if needed
JP1_df <- read.xlsx(xlsxFile = 'japanP1.xlsx')

```

```{r}
str(JP1_df)
summary(JP1_df)
```
MS = market spend %
```{r}
pairs(JP1_df)
GGally::ggpairs(JP1_df)
```

# Fitting the linear model
```{r}
lm_fit = lm(MktShr ~ . - RQ, JP1_df)
```

```{r}
summary(lm_fit)
```

```{r}
resid_fitted <- ggplot(data = lm_fit, aes(x= .fitted, y= .resid)) + geom_point() + labs(x= "Fitted Values" , y= "Residuals")
resid_fitted
```

```{r}
plot(JP1_df$MyG, resid(lm_fit))
plot(JP1_df$AvgG, resid(lm_fit))
plot(JP1_df$RP, resid(lm_fit))
plot(JP1_df$RQ, resid(lm_fit))
plot(JP1_df$MS, resid(lm_fit))
plot(JP1_df$price, resid(lm_fit))
plot(JP1_df$LQ1, resid(lm_fit))
plot(JP1_df$LQ2, resid(lm_fit))
plot(JP1_df$LQ3, resid(lm_fit))
plot(JP1_df$mg1, resid(lm_fit))
plot(JP1_df$mg2, resid(lm_fit))
plot(JP1_df$mg3, resid(lm_fit))
```