---
title: "Customer Preferences"
author: "Steven Barnard"
date: "September 7, 2018"
output:
  pdf_document: null
  number_sections: yes
  html_document:
    df_print: paged
  toc: yes
  toc_depth: 4
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\tableofcontents 
\newpage

### Load the needed libraries

```{r load libraries, message=FALSE, warning=FALSE}

rm(list=ls())

if (!require('devtools')) install.packages('devtools')
library('devtools')

if (!require('openxlsx')) install.packages('openxlsx')
library('openxlsx')

if (!require('ggplot2')) install.packages('ggplot2')
library('ggplot2')

if (!require('dplyr')) install.packages('dplyr')
library('dplyr')

if (!require('glmnet')) install.packages('glmnet')
library('glmnet')

if (!require('data.table')) install.packages('data.table')
library('data.table')

if (!require('plotmo')) install.packages('plotmo')
library('plotmo')

if (!require('GGally')) install.packages('GGally')
library('GGally')
```

### Load the data into R

This data is trying to understand the customer preferences for the management game simulation.      

```{r Read the data, echo=TRUE }
SimData_df <- read.xlsx(xlsxFile = "CustPref_RawData(R).xlsx", sheet = 'MkRawData')
SimData_df <- as_data_frame(SimData_df)
```

###  Reviewing structure of source data  

```{r Look at the structure of the data.frame}
#str(SimData_df)
```

###  Transform the data as needed


```{r data transformation section}

SimData_df$Country = as.factor(SimData_df$Country)
#HW1_df <-mutate(HW1_df, X4_2 = X4*X4)
#str(SimData_df)
```

###  Segregate Data by Country and by Product before running regressions


```{r data segregation}

JP1_df <- filter(SimData_df,Country == "Japan", Product == "1", Time.Period >= 4)
MX1_df <- filter(SimData_df,Country == "Mexico", Product == "1", Time.Period >= 4)
CN1_df <- filter(SimData_df,Country == "China", Product == "1", Time.Period >= 4)
UK1_df <- filter(SimData_df,Country == "U.K.", Product == "1", Time.Period >= 4)
GR1_df <- filter(SimData_df,Country == "Germany", Product == "1", Time.Period >= 4)
US1_df <- filter(SimData_df,Country == "U.S.", Product == "1", Time.Period >= 4)

JP2_df <- filter(SimData_df,Country == "Japan", Product == "2", Time.Period >= 4)
MX2_df <- filter(SimData_df,Country == "Mexico", Product == "2", Time.Period >= 4)
CN2_df <- filter(SimData_df,Country == "China", Product == "2", Time.Period >= 4)
UK2_df <- filter(SimData_df,Country == "U.K.", Product == "2", Time.Period >= 4)
GR2_df <- filter(SimData_df,Country == "Germany", Product == "2", Time.Period >= 4)
US2_df <- filter(SimData_df,Country == "U.S.", Product == "2", Time.Period >= 4)

```
### Learn about the data set with Summary statistics and plots

Summary statistics review for the variables to get familiarized with the data. 

```{r summary statistics}
#summary(JP1_df)

```

### Reviewing general relationships in the entire data set.  Using "pairs"" function from the base graphics package and "GGally::ggpairs"" from the GGally and ggplot2 packages. 

```{r pairs summary of the data}
#pairs(JP1Data)
```

```{r GGally_ggpairs, message=FALSE, warning=FALSE}
#GGally::ggpairs(JP1Data)
```

### Fit a Linear Regression model explain the annual charges

Now fit the linear regression model using the lm function.  

```{r fitting the linear regression}



```

# JAPAN_1
```{r summarize JP1_fit output}
JP1_fit =lm(MktShare ~ . , data = select(JP1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(JP1_fit)
JP1_fit =lm(MktShare ~ . -L2_RelQuality -RelQuality -L3_RelGreen -RelGreen -L3_RelQuality -L3_RelBrandMkt -L2_RelGreen, 
            data = select(JP1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(JP1_fit)

#ggplot(JP1_fit, aes(x = RelPrice, y = MktShare)) + geom_point() + stat_smooth(method = "lm", col = "red")
#ols.coef = JP1_df$coefficients
```
# MEXICO_1
```{r summarize JP1_fit output}
MX1_fit =lm(MktShare ~ . , data = select(MX1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(MX1_fit)
MX1_fit =lm(MktShare ~ . -L3_RelGreen -RelGreen -RelQuality -L2_RelQuality -L3_RelQuality -L2_RelGreen -L3_RelBrandMkt, 
            data = select(MX1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(MX1_fit)
```
# CHINA_1
```{r summarize JP1_fit output}
CN1_fit =lm(MktShare ~ . , data = select(CN1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(CN1_fit)
CN1_fit =lm(MktShare ~ . -L3_RelBrandMkt -L3_RelBrandMkt -L2_RelQuality -L3_RelGreen -L3_RelQuality -L2_RelGreen -RelGreen -RelQuality, 
            data = select(CN1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(CN1_fit)
```
# UK_1
```{r summarize JP1_fit output}
UK1_fit =lm(MktShare ~ . , data = select(UK1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(UK1_fit)
UK1_fit =lm(MktShare ~ . -L2_RelGreen -L3_RelQuality -RelGreen -L2_RelQuality -L3_RelBrandMkt -RelQuality -L3_RelGreen, 
            data = select(UK1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(UK1_fit)
```

# GR_1
```{r summarize JP1_fit output}
GR1_fit =lm(MktShare ~ . , data = select(GR1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(GR1_fit)
GR1_fit =lm(MktShare ~ . -L2_RelGreen -RelQuality -RelGreen -L3_RelQuality -L3_RelBrandMkt -L2_RelQuality -L3_RelGreen, 
            data = select(GR1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(GR1_fit)
```

# US_1
```{r summarize JP1_fit output}
US1_fit =lm(MktShare ~ . , data = select(US1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(US1_fit)
US1_fit =lm(MktShare ~ . -L2_RelGreen -RelQuality -L2_RelQuality -RelGreen -L3_RelQuality, 
            data = select(US1_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(US1_fit)
```

# JAPAN_2
```{r summarize JP1_fit output}
JP2_fit =lm(MktShare ~ . , data = select(JP2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(JP2_fit)
JP2_fit =lm(MktShare ~ . -L1_RelBrandMkt -L1_RelQuality -L2_RelQuality -L3_RelQuality -L3_RelGreen -L2_RelGreen -RelGreen -L3_RelBrandMkt, 
            data = select(JP2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(JP2_fit)
```

# MEXICO_2
```{r summarize JP1_fit output}
MX2_fit =lm(MktShare ~ . , data = select(MX2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(MX2_fit)
MX2_fit =lm(MktShare ~ . -L3_RelBrandMkt -L2_RelGreen -L2_RelQuality -L3_RelGreen -RelGreen -RelQuality -L3_RelQuality, 
            data = select(MX2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(MX2_fit)
```

# CHINA_2
```{r summarize JP1_fit output}
CN2_fit =lm(MktShare ~ . , data = select(CN2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(CN2_fit)
CN2_fit =lm(MktShare ~ . -L1_RelQuality -L2_RelGreen -L2_RelQuality -L3_RelGreen -RelGreen -L3_RelQuality -L3_RelBrandMkt, 
            data = select(CN2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(CN2_fit)
```

# UK_2
```{r summarize JP1_fit output}
UK2_fit =lm(MktShare ~ . , data = select(UK2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(UK2_fit)
UK2_fit =lm(MktShare ~ .  -RelQuality -L2_RelQuality -L3_RelQuality -L2_RelGreen -RelGreen -L3_RelGreen -L3_RelBrandMkt, 
            data = select(UK2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(UK2_fit)
```

# GR_2
```{r summarize JP1_fit output}
GR2_fit =lm(MktShare ~ . , data = select(GR2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(GR2_fit)
GR2_fit =lm(MktShare ~ . -L2_RelQuality -L2_RelGreen -L3_RelQuality -RelGreen -RelQuality, 
            data = select(GR2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(GR2_fit)
```

# US_2
```{r summarize JP1_fit output}
US2_fit =lm(MktShare ~ . , data = select(US2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(US2_fit)
US2_fit =lm(MktShare ~ . -RelQuality -L3_RelQuality -L3_RelGreen -L2_RelQuality -L3_RelBrandMkt, 
            data = select(US2_df, -Country, -Product, -Time.Period, -AbsGreen, -AvgGreen, -RelBrandMkt))
summary(US2_fit)
```

#### Residuals Histogram 

First consider the histogram plot of the residuals.  This is to help judge if the errors are normally distributed.
```{r write to file}

write.xlsx(JP1_fit, file = "JP1_fit.xlsx", sheetName = "JP1_fit", row.names=FALSE)
write.xlsx(MX1_fit, file = "MX1_fit.xlsx", sheetName = "MX1_fit", append=TRUE, row.names=FALSE)
write.xlsx(CN1_fit, file = "CN1_fit.xlsx", sheetName = "CN1_fit", append=TRUE, row.names=FALSE)
write.xlsx(UK1_fit, file = "UK1_fit.xlsx", sheetName = "UK1_fit", append=TRUE, row.names=FALSE)
write.xlsx(GR1_fit, file = "GR1_fit.xlsx", sheetName = "GR1_fit", append=TRUE, row.names=FALSE)
write.xlsx(US1_fit, file = "US1_fit.xlsx", sheetName = "US1_fit", append=TRUE, row.names=FALSE)

write.xlsx(JP2_fit, file = "JP2_fit.xlsx", sheetName = "JP2_fit", append=TRUE, row.names=FALSE)
write.xlsx(MX2_fit, file = "MX2_fit.xlsx", sheetName = "JMX_fit", append=TRUE, row.names=FALSE)
write.xlsx(CN2_fit, file = "CN2_fit.xlsx", sheetName = "CN2_fit", append=TRUE, row.names=FALSE)
write.xlsx(UK2_fit, file = "UK2_fit.xlsx", sheetName = "UK2_fit", append=TRUE, row.names=FALSE)
write.xlsx(GR2_fit, file = "GR2_fit.xlsx", sheetName = "GR2_fit", append=TRUE, row.names=FALSE)
write.xlsx(US2_fit, file = "US2_fit.xlsx", sheetName = "US2_fit", append=TRUE, row.names=FALSE)
```

### Diagnostic plots to judge the model's fit

Now consider some diagnostic to judge how well the model summarizing the data.

#### Residuals Histogram 

First consider the histogram plot of the residuals.  This is to help judge if the errors are normally distributed.
```{r residual histogram, message=FALSE, warning=FALSE}
#resid_hist <- ggplot(data = lm_fit, aes(x=.resid)) + geom_histogram()
#resid_hist
```

Note the long upper tail.  This suggests  nonnormal errors. 

#### Residuals qq-plot

Next consider the qq-plot to again judge the normality of the errors. 

```{r qq-plot of the residuals}
#resid_qqplot <- ggplot(dat = lm_fit, aes(sample=.stdresid)) + stat_qq() + geom_abline()
#resid_qqplot
```
The qq-plot is not linear.  Again, this demonstrates the errors are not normally distributed.  

#### Residuals versus Fitted scatterplot

Finally, plot the residuals against the fitted values to see if there is additional structure in the not captured by the model. 

```{r rediduals versus fitted}
#resid_fitted <- ggplot(data = lm_fit, aes(x=.fitted, y=.resid)) + geom_point() +
#             labs(x="Fitted Values", y="Residuals")
#resid_fitted

```
